---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Get TB Data in R

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/getTBinR)](https://cran.r-project.org/package=getTBinR)
[![Travis-CI Build Status](https://travis-ci.org/<USERNAME>/<REPO>.svg?branch=master)](https://travis-ci.org/<USERNAME>/<REPO>)
[![Coverage Status](https://img.shields.io/codecov/c/github/<USERNAME>/<REPO>/master.svg)](https://codecov.io/github/<USERNAME>/<REPO>?branch=master)

This package contains functions to download and clean data made available by the World Health Organisation (WHO), see [here](http://www.who.int/about/copyright/en/) for permissions.

## Installation

You can install getTBinR from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("seabbs/getTBinR")
```

### Docker

This package has been developed in docker based on the `rocker/tidyverse` image, to access the  development environment enter the following at the command line (with an active docker daemon running),

```{bash, eval = FALSE}
docker pull seabbs/gettbinr
docker run -d -p 8787:8787 -p 54321:54321 -e USER=getTBinR -e PASSWORD=gettbinr --name getTBinR seabbs/gettbinr
```

The rstudio client can be accessed on port `8787` at `localhost` (or your machines ip). The default username is getTBinR and the default password is getTBinR.
