---
title: "Tuberculosis Report"
output: html_document
params: 
  country: "United Kingdom"
  interactive: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warnings = FALSE,
                      eval = TRUE)
```

```{r report-setup, include = FALSE, results = "hide"}
## Load the package
library(getTBinR)

## Load additional packages
library(dplyr) # For data munging 
library(tidyr)
library(rlang)
library(ggplot2)

## Get the data
tb <- get_tb_burden(verbose = FALSE)

## Get the data dictionary
dict <- get_data_dict(verbose = FALSE)

##Assign parameters
country <- params$country
interactive <- params$interactive
```

## TB incidence rates

```{r get country-stats, include = FALSE, results = "hide"}
metric_summary <- function(df = NULL, target_country = NULL, metric = NULL) {
  
  target_country <- df$country[grepl(target_country, df$country)] %>% 
    unique %>% 
    first
    
  ##  Set up metric with confidence intervals
  metric <- enquo(metric)
  metric_lo <- sym(paste0(quo_name(metric), "_lo"))
  metric_hi <- sym(paste0(quo_name(metric), "_hi"))
  
  ## Filter for the country of interest
  country_df <- df %>% 
  filter(country %in% target_country)
  
## Most up to date year of incidence data
recent_inc <- country_df %>% 
  drop_na(!!metric) %>% 
  filter(year == max(year)) %>% 
  select(!!metric, !!metric_lo, !!metric_hi, year, g_whoregion) %>% 
  mutate(inc_rate = paste0(!!metric, " (", !!metric_lo, " - ", !!metric_hi, ")"))

## Country rank
ranked_countries_inc <- df %>% 
  filter(year == recent_inc$year) %>% 
  arrange(desc(!!metric)) %>% 
  mutate(rank = 1:n())

## World rank
target_rank_world <- ranked_countries_inc %>% 
  filter(country == target_country) %>% 
  pull(rank)

## Region rank
target_rank_region <- ranked_countries_inc %>% 
  filter(g_whoregion %in% recent_inc$g_whoregion) %>% 
  mutate(rank = 1:n()) %>% 
  filter(country == target_country) %>%
  pull(rank)


## Summarise annual change
country_change <- summarise_tb_burden(metric = quo_name(metric),
                                      stat = "mean",
                                      countries = target_country,
                                      compare_to_region = FALSE,
                                      compare_to_world = FALSE,
                                      compare_all_regions = FALSE,
                                      annual_change = TRUE,
                                      verbose = FALSE) %>% 
  filter(year > (max(year) - 10)) %>% 
  summarise(change = mean(!!metric, na.rm = FALSE)) %>%
  mutate(change = round(change * 100, 1) %>% 
           paste0(., "%")) %>% 
  pull(change)

out <- list(recent_inc$year[1], recent_inc$inc_rate[1],
            target_rank_world, target_rank_region,
            country_change)

names(out) <- c("year", "metric", "world_rank", "region_rank", "avg_change")

out <- ifelse(is.na(out), "(Missing)", out)
return(out)
}
  
  
inc_sum <- metric_summary(tb, country, e_inc_100k)
```

In `r inc_sum$year` `r country` had an estimated Tuberculosis incidence rate of `r inc_sum$metric` per 100,000 people making it number `r inc_sum$world_rank` in the world and number `r inc_sum$region_rank` regionally. In the last 10 years this has changed by `r inc_sum$avg_change` on average each year. 

### Regional and Global Trends Comparision

```{r}
plot_tb_burden_summary(countries = country,
                       metric_label = "e_inc_100k",
                       compare_to_world = TRUE, 
                       compare_to_region = TRUE,
                       compare_all_regions = FALSE,
                       annual_change = FALSE,
                       facet = "Area",
                       scales = "free_y",
                       legend = "none",
                       interactive = interactive,
                       verbose = FALSE)
```

## Rates Regional Breakdown

```{r, fig.height = 8, fig.width = 8 }
plot_tb_burden_overview(countries = country,
                        compare_to_region = TRUE,
                        interactive = interactive,
                        verbose = FALSE)
```

## Case Detection Rates (CDR)

```{r, include = FALSE, results = "hide"}
cdr_sum <- metric_summary(tb, country, c_cdr)
```

`r country` had an estimated case detection rate of `r cdr_sum$metric`% in `r cdr_sum$year` making it number `r cdr_sum$world_rank` in the world (with number 1 having the highest CDR) and number `r cdr_sum$region_rank` regionally. In the last 10 years this has changed by `r cdr_sum$avg_change` on average each year. 

### Regional Breakdown

```{r, fig.height = 8, fig.width = 8}
plot_tb_burden_overview(metric = "c_cdr",
                        countries = country,
                        compare_to_region = TRUE,
                        interactrive = interactive,
                        verbose = FALSE)
```

## TB mortality rates - excluding HIV


```{r, include = FALSE, results = "hide"}
mort_exc_hiv_sum <- metric_summary(tb, country, e_mort_exc_tbhiv_100k)
```

In `r mort_exc_hiv_sum$year` `r country` had an estimated Tuberculosis mortality rate (excluding HIV) of `r mort_exc_hiv_sum$metric`  per 100,000 people making it number `r mort_exc_hiv_sum$world_rank` in the world and number `r mort_exc_hiv_sum$region_rank` regionally. In the last 10 years this has changed by `r mort_exc_hiv_sum$avg_change` on average each year. 

### Proportion of TB Cases that Died (excluding HIV) - Regional and Global Comparision

```{r, fig.height = 8, fig.width = 12}
plot_tb_burden_summary(metric = "e_mort_exc_tbhiv_num",
                       denom = "e_inc_num",
                       rate_scale = 100,
                       countries = country,
                       compare_to_region = TRUE,
                       compare_all_regions = FALSE,
                       interactive = interactive,
                       verbose = FALSE,
                       facet = "Area",
                       scales = "free_y",
                       legend = "none") +
  labs(y = "Proportion (%) of TB cases that died (excluding HIV)")
```

### Rates Regional Breakdown

```{r, fig.height = 8, fig.width = 8}
plot_tb_burden_overview(metric = "e_mort_exc_tbhiv_100k",
                        countries = country,
                        compare_to_region = TRUE,
                        interactrive = interactive,
                        verbose = FALSE)
```

## TB HIV related mortality rates

```{r, include = FALSE, results = "hide"}
mort_inc_hiv_sum <- metric_summary(tb, country, e_mort_tbhiv_100k)
```

In `r mort_inc_hiv_sum$year` `r country` had an estimated Tuberculosis mortality rate (related to HIV) of `r mort_inc_hiv_sum$metric`  per 100,000 people making it number `r mort_inc_hiv_sum$world_rank` in the world and number `r mort_inc_hiv_sum$region_rank` regionally. In the last 10 years this has changed by `r mort_inc_hiv_sum$avg_change` on average each year. 

### Proportion of TB Cases that Died (related to HIV) - Regional and Global Comparision

```{r, fig.height = 8, fig.width = 12}
plot_tb_burden_summary(metric = "e_mort_tbhiv_num",
                       denom = "e_inc_num",
                       rate_scale = 100,
                       countries = country,
                       compare_to_region = TRUE,
                       compare_all_regions = FALSE,
                       interactive = interactive,
                       verbose = FALSE,
                       facet = "Area",
                       scales = "free_y",
                       legend = "none") +
  labs(y = "Proportion (%) of TB cases that died (related to HIV)")
```

### Rates Regional Breakdown

```{r, fig.height = 8, fig.width = 8}
plot_tb_burden_overview(metric = "e_mort_tbhiv_100k",
                        countries = country,
                        compare_to_region = TRUE,
                        interactrive = interactive,
                        verbose = FALSE)

```
