---
title: "Getting started with getTBinR"
author: "Sam Abbott"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with getTBinR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 7, 
  fig.align = "center"
)
```

## Using the package

First load the package. We also load several other packages to help quickly explore the data.

```{r packages, message = FALSE}
library(getTBinR)
library(ggplot2)
library(knitr)
library(magrittr)
library(dplyr)
```


## Getting TB burden data

Get TB burden data with a single function call. This will download the data if it has never been accessed and then save a local copy to R's temporary directory (see `tempdir()`). If a local copy exists from the current session then this will be loaded instead.

```{r get-tb-burden-data}
tb_burden <- get_tb_burden()

tb_burden
```

## Searching for variable definitions

The WHO provides a large, detailed, data dictionary for use with the TB burden data. However, searching through this dataset can be tedious. To streamline this process `getTBinR` provides a search function to find the definition of a single or multiple variables. Again if not previously used this function will download the data dictionary to the temporary directory, but in subsequent uses will load a local copy.

```{r search-data-dict}
vars_of_interest <- search_data_dict(var = c("country",
                                             "e_inc_100k",
                                             "e_inc_100k_lo",
                                             "e_inc_100k_hi"))

knitr::kable(vars_of_interest)
```

We might also want to search the variable definitions for key phrases, for example mortality.

```{r search-data-defs}
defs_of_interest <- search_data_dict(def = c("mortality"))

knitr::kable(defs_of_interest)
```

Finally we could both search for a known variable and for key phrases in variable definitions.

```{r search-data-defs-var}
vars_defs_of_interest <- search_data_dict(var = c("country"),
                                     def = c("mortality"))

knitr::kable(vars_defs_of_interest)
```

## Mapping Global Incidence Rates

To start exploring the WHO TB data we map, the most recently available, global TB incidence rates. Mapping data can help identify spatial patterns.

```{r map-tb-incidence-eur, fig.width = 15}
getTBinR::map_tb_burden(metric = "e_inc_100k")
```

## Plotting Incidence Rates for All Countries

To showcase how quickly we can go from no data to plotting informative graphs we quickly explore incidence rates for all countries in the WHO data.


```{r plot-tb-all, fig.height=20, fig.width = 10}
getTBinR::plot_tb_burden_overview(metric = "e_inc_100k",
                                  interactive = FALSE)
```


Another way to compare incidence rates in countries is to look at the annual percentage change. The plot below only shows countries with a maximum incidence rate above 5 per 100,000.


```{r plot-tb-al-per, fig.height=20, fig.width = 10}
higher_burden_countries <- tb_burden %>% 
  group_by(country) %>% 
  summarise(e_inc_100k = min(e_inc_100k)) %>% 
  filter(e_inc_100k > 5) %>% 
  pull(country) %>% 
  unique

getTBinR::plot_tb_burden_overview(metric = "e_inc_100k",
                                  interactive = FALSE,
                                  annual_change = TRUE,
                                  countries = higher_burden_countries)
```

## Summarising Regional and Global Incidence Rates

We might also be interested in getting a regional/global overview of TB incidence rates (*Hint: Use `search_data_dict` to look up `e_inc_100k` to see what role this is playing here*). See `?plot_tb_burden_summary` for more ways to summarise TB metrics.

```{r plot-regional-summary} 
getTBinR::plot_tb_burden_summary(conf = NULL, metric_label = "e_inc_100k")
```

## Plotting Incidence Rates over Time in 9 Randomly Sampled Countries

Diving deeper into the data lets plot a sample of 9 countries using the inbuilt `plot_tb_burden` function. Again plotting incidence rates, but this time with 95% confidence intervals. As you can see this isn't a hugely informative graph. Lets improve it!

```{r plot-incidence}
## Take a random sample of countries
sample_countries <- sample(unique(tb_burden$country), 9)
plot_tb_burden(tb_burden, metric = "e_inc_100k",
               countries = sample_countries,
               legend = "top")
```

We have faceted by country so that we can more easily see what is going on. This allows us to easily explore between country variation - depending on the sample there is likely to be alot of this.

```{r plot-incidence-facet}
plot_tb_burden(tb_burden, metric = "e_inc_100k",
               countries = sample_countries,
               facet = "country")
```

To explore within country variation we need to change the scale of the y axis.

```{r plot-incidence-facet-free-y}
plot_tb_burden(tb_burden, metric = "e_inc_100k",
               countries = sample_countries,
               facet = "country",
               scales = "free_y")
```

We might also be interested in mortality in both HIV negative and HIV positive cases in our sample countries. We can also look at this using `plot_tb_burden` as follows. Note we can do this without specifying the TB burden data, the plotting function will automatically find it either locally or remotely.

```{r plot-mortality-excluding-hiv}
plot_tb_burden(metric = "e_mort_exc_tbhiv_100k",
               countries = sample_countries,
               facet = "country",
               scales = "free_y")
```

```{r plot-mortality-including-hiv}
plot_tb_burden(metric = "e_mort_tbhiv_100k",
               countries = sample_countries,
               facet = "country",
               scales = "free_y")
```

